---
title: "Prepare building footprints and heights for Liverpool"
output: html_document
---

## Load packages

```{r, echo=FALSE}
library(tidyverse)
library(vroom)
library(sf)
```

## Set root directory

Set path to your wd in masterfile or directly here. 

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(my_wd))
```

## Load Data

* Case study region
* Building footprints 2011 and 2021: OS Mastermap data comes as tiles which need to be merged. We extract buildings only. 
* Building heights 2021

Details on Building heights: 
There are different ways of measuring building height in the UK data. For further processing, we use RelHMax and RelH2. 

* RelH2 = AbsH2 – AbsHMin. RelH2 is the relative building height from the base of the building (AbsHMin) to the height of the base of the roof (AbsH2).
* RelHMax = AbsHMax – AbsHMin. RelHMax is the relative building height from the base of the building (AbsHMin) to the height of the highest point on the building (AbsHMax).

See details in: https://docs.os.uk/os-downloads/topography/os-mastermap-topography-layer-building-height-attribute/os-mastermap-topography-layer-building-height-attribute-technical-specification/feature-type/building-height-attribute

```{r}
liv_case <- st_read("./output/liv_iso30_LSOA_dissolved.gpkg")

liv_bldg_2021 <-
  list.files(".input/OS/Liverpool 2021", full.names = TRUE) %>% 
  str_subset("gpkg$") %>% 
  map_dfr(read_sf, layer = "Topographicarea") %>% 
  mutate(fid_os = fid) %>% select(-fid) %>%
  filter(theme == "Buildings")

liv_bldg_2011 <-
  list.files(".input/OS/Liverpool 2011", full.names = TRUE) %>% 
  str_subset("gpkg$") %>% 
  map_dfr(read_sf, layer = "Topographicarea") %>% 
  mutate(fid_os = fid) %>% select(-fid) %>%
  filter(theme == "Buildings")

liv_2021_ha <-
  list.files(".input/OS/building_heights/Liverpool Oct 2021/", recursive = TRUE, 
             full.names = TRUE) %>% 
  str_subset("(?<=\\/)[a-zA-Z0-9]{6,6}\\.csv$") %>%
  map_dfr(read_csv, col_names = FALSE) 

colnames(liv_2021_ha) <- c("OS_TOPO_TOID", "OS_TOPO_TOID_VERSION", "BHA_ProcessDate", "TileRef", "AbsHMin", "AbsH2", "AbsHMax", "RelH2", "RelHMax", "BHA_Conf")
```

## Clean data heights

As the data comes as tiles, there are some building polygons duplicated at the intersections. We filter them. 

```{r}
liv_2021_ha_clean <- liv_2021_ha %>%
  group_by(OS_TOPO_TOID) %>% 
  filter(!duplicated(OS_TOPO_TOID))
```

## Join building footprints and buildings heights 2021

```{r}
liv_bldg_2021_height <- liv_bldg_2021 %>%
  left_join(liv_2021_ha_clean, by = join_by("fid_os" == "OS_TOPO_TOID"))
```

## Filter to buildings which intersect case study region

```{r}
st_crs(liv_case) == st_crs(liv_bldg_2021_height)
st_crs(liv_case) == st_crs(liv_bldg_2011)

liv_bldg_2021_height_case <- st_filter(liv_bldg_2021_height, liv_case, .predicate = st_intersects)

liv_2011_case <- st_filter(liv_bldg_2011, liv_case, .predicate = st_intersects)
```

```{r}
liv_bldg_2021_clean <- liv_bldg_2021_height_case %>%
  group_by(fid_os) %>% 
  filter(!duplicated(fid_os)) 

liv_bldg_2011_clean <- liv_2011_case %>%
  group_by(fid_os) %>% 
  filter(!duplicated(fid_os)) 
```


## Write results

```{r}
st_write(liv_bldg_2021_clean, "./output/liv_bldg_2021_01_ha_raw.gpkg", delete_layer = TRUE)
st_write(liv_bldg_2011_clean, "./output/liv_bldg_2011_01_raw.gpkg", delete_layer = TRUE)
```
