---
title: "Derive densification types"
output: html_document
---

## Load packages

```{r, echo=FALSE}
library(tidyverse)
library(dplyr)
library(sf)
library(uuid)
```

## Set root directory

Set path to your wd in masterfile or directly here. 

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(my_wd))
```

## Load Data

* Results of the building evolution process (Output files of the python process need to be copied manually to output-folder of the project.)

```{r}
bldg_evol_str <- st_read("./output/compdense_str_EVOLUTION.gpkg")
bldg_evol_dor <- st_read("./output/compdense_dor_EVOLUTION.gpkg")
bldg_evol_liv <- st_read("./output/compdense_liv_EVOLUTION.gpkg")
```

## Reduce complex output of building evolution to actual densification

The building evolution process identifies every type of building change, like appeared, disappeared, split, merged, complex change. The latter are changes which are most likely changes in how the footprints were modeled. In reality, this does not mean that a building was e.g. subdivided in several housing units. Thus, we do not count them as densification. Only buildings which are labelled as appeared are considered relevant to measure densification. We compare them with disappeared buildings to understand if they are the result of reconstruction or infill. 

```{r}
derive_denstype <- function(bldg_evol_region) {
  
  ## Step 1: Add UUID field
  bldg_evol_region <- bldg_evol_region %>%
    mutate(id_univ = paste0("bldg_", row_number()))
  
  ## Step 2: Filter disappeared and appeared features
  disappeared <- bldg_evol_region %>% filter(type == "disappeared")
  appeared <- bldg_evol_region %>% filter(type == "appeared")
  
  ## Step 3: Identify appeared features touching disappeared ones
  touched <- st_intersects(appeared, disappeared, sparse = FALSE)
  appeared <- appeared %>%
    mutate(reconstruction = apply(touched, 1, any)) %>%
    st_drop_geometry()
  
  ## Step 4: Join reconstruction attribute back to full dataset
  bldg_with_reconstruction <- left_join(bldg_evol_region,
                                        appeared %>% select(id_univ, reconstruction),
                                        by = "id_univ")
  
  ## Step 5: Derive densification type
  bldg_with_densification <- bldg_with_reconstruction %>%
    mutate(
      denstype = case_when(
        type == "appeared" & reconstruction == TRUE ~ "reconstruction",
        type == "appeared" & reconstruction == FALSE ~ "infill",
        type == "disappeared" ~ "disappeared",
        TRUE ~ NA_character_
      ),
      denstype = if_else(is.na(denstype), "stable", denstype)
    )
  
  bldg_with_densification
  
}
```

```{r}
str_denstype <- derive_denstype(bldg_evol_str)
dor_denstype <- derive_denstype(bldg_evol_dor)
liv_denstype <- derive_denstype(bldg_evol_liv)
```

## Write results

```{r}
st_write(str_denstype, "./output/str_bldg_densetype.gpkg", delete_layer = TRUE)
st_write(dor_denstype, "./output/dor_bldg_densetype.gpkg", delete_layer = TRUE)
st_write(liv_denstype, "./output/liv_bldg_densetype.gpkg", delete_layer = TRUE)
```
