---
title: "Building type Liverpool"
output: html_document
---

## Load packages

```{r, echo=FALSE}
library(sf)
library(dplyr)
library(uuid)
library(stringr)
```

## Set root directory

Set path to your wd in masterfile or directly here. 

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(my_wd))
```

## Load Data

* AddressBase Premium: https://www.ordnancesurvey.co.uk/products/addressbase-premium (license needed)
   + BLPU: Basic land and property unit
   + Classification: This record holds the classification of a property and allows one to search upon the use of a feature.
* Building footprints 2021

```{r}
# Basic Land and Property Unit
addb_blpu <- st_read("./input/OS/AddressBase/0040176195-6425786-1/6425786.gpkg/6425786.gpkg", layer = "blpu")
addb_clas <- st_read("./input/OS/AddressBase/0040176195-6425786-1/6425786.gpkg/6425786.gpkg", layer = "classification")

# ## for faster processing only load the joined data after it was once saved
# addressbase <- st_read("./output/liv_address_base_01_all.gpkg")

liv_bldg_2021 <- st_read("./output/liv_bldg_2021_02_filtered.gpkg")
```

## Join address_base data

For building classification we need to join the data from blpu with classification. 

```{r}
addressbase <- addb_blpu %>% 
  select(uprn, start_date, end_date) %>% 
  left_join(addb_clas %>% 
              transmute(uprn, class_key, classification_code, 
                        start_date_clas = start_date, 
                        end_date_clas = end_date))
```

## Filter to relevant addresses

We want to identify residential addresses. The dataset includes addresses throughout the property lifecycle, from proposed planning development through to demolished address. Thus, we need to filter to properties which were created prior to 2021 and still existed. Further, we filter to classification codes which are linked to residential use. 

AddressBase Classification scheme: https://docs.os.uk/os-downloads/addressing-and-location/addressbase-fundamentals/classification-scheme

```{r}
addressbase_res <- addressbase %>% 
  filter(start_date_clas <= "2021-06-01") %>% # filter to addresses which were created prior to 2021
  filter(end_date_clas >= "2021-06-01" |is.na(end_date_clas)) %>%  # filter to addresses which still existed in 2021
  filter(classification_code %in% c("RD02", # detached
                                    "RD03", # semi-detached
                                    "RD04", # terraced
                                    "RD06", # self-contained flat
                                    "RH01", # house in multiple occupation
                                    "X"))   # dual use
```

## Intersect buildings with address data

```{r}
## join address_information to buildings
joined <- st_join(liv_bldg_2021, addressbase_res, join = st_intersects, left = TRUE)
joined_reduced <- joined %>% filter(!is.na(classification_code)) %>% 
  st_drop_geometry() %>% 
  group_by(fid_os) %>% 
  summarise(n_address = n(), 
            classification_code = paste(sort(unique(classification_code)) , collapse = ", "))

# aggregate to building types
bldg_type <- joined_reduced %>% 
  mutate(sfh_mfh = if_else(str_detect(classification_code, "RD06|RH01|RH03"), "mfh", "other"), 
         sfh_mfh = if_else(str_detect(classification_code, "RD02|RD03|RD04") & !str_detect(classification_code, "RD06|RH01|RH03"), "sfh", sfh_mfh), 
         sfh_mfh = if_else(str_detect(classification_code, "X"), "mfh", sfh_mfh))

## final dataset
liv_bldg_2021_addbase <- liv_bldg_2021 %>% 
  left_join(bldg_type) %>% 
  mutate(res_non_res = if_else(is.na(sfh_mfh), "non_res", "res"))
```

## Write results

```{r}
st_write(liv_bldg_2021_addbase, "./output/liv_bldg_2021_03_bldg_type.gpkg", delete_layer = TRUE)

# st_write(addressbase, "./output/liv_address_base_01_all.gpkg", delete_layer = TRUE)
# st_write(addressbase_res, "./output/liv_address_base_02_reduced.gpkg", delete_layer = TRUE)
```

