---
title: "Prepare Corine land cover"
output: html_document
---

## Load packages

```{r, echo=FALSE}
library(dplyr)
library(sf)
```

## Set root directory

Set path to your wd in masterfile or directly here. 

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(my_wd))
```

## Load Data

* Case study region
* Download CLC 2012 data from https://land.copernicus.eu/en/products/corine-land-cover/clc-2012 (free account needed)

As loading the full dataset runs very long, we ruffly reduced the dataset to our case study regions in QGIS. 

```{r}
dor_case <- st_read("./output/dor_iso30_VWG_dissolved.gpkg")
str_case <- st_read("./output/str_iso30_commune_dissolved.gpkg")
liv_case <- st_read("./output/liv_iso30_LSOA_dissolved.gpkg")

dor_clc_raw <- st_read("./input/dor_orig_corine_u2018_clc2012_20u1.gpkg")
str_clc_raw <- st_read("./input/str_orig_corine_u2018_clc2012_20u1.gpkg")
liv_clc_raw <- st_read("./input/liv_orig_corine_u2018_clc2012_20u1.gpkg")
```

## Prepare data and crop to case study regions

We are interested in urban land use. See nomenclature here: https://land.copernicus.eu/content/corine-land-cover-nomenclature-guidelines/html/

```{r}
prepare_clc <- function(region_clc_raw, region_case) {
  
  region_clc <- region_clc_raw %>%
    mutate(
      code_simple = substr(Code_12, start = 1, stop = 2),
      code_nam = if_else(code_simple == "11", "urban fabric", "x"),
      code_nam = if_else(code_simple == "12", "industrial/comercial/transport", code_nam),
      code_nam = if_else(code_simple == "13", "mine/dump/construction sites", code_nam),
      code_nam = if_else(code_simple == "14", "green ua/sport/leisure", code_nam)
    ) %>%
    filter(code_simple %in% c("11", "12", "13", "14"))
  
  region_clc <- st_transform(region_clc, st_crs(region_case))
  region_clc_case <- st_intersection(region_clc, region_case)
  
  region_clc_case
}
```

```{r}
str_clc_case <- prepare_clc(str_clc_raw, str_case)
dor_clc_case <- prepare_clc(dor_clc_raw, dor_case)
liv_clc_case <- prepare_clc(liv_clc_raw, liv_case)
```

## Prepare urban mask 2012

```{r}
str_clc_urban_mask <- str_clc_case %>% 
  group_by() %>%
  summarise(geom = st_union(geom)) %>% 
  mutate(clc_inside = "str_clc_inside")

dor_clc_urban_mask <- dor_clc_case %>% 
  group_by() %>%
  summarise(geom = st_union(geom)) %>% 
  mutate(clc_inside = "dor_clc_inside")

liv_clc_urban_mask <- liv_clc_case %>% 
  group_by() %>%
  summarise(geom = st_union(geom)) %>% 
  mutate(clc_inside = "liv_clc_inside")
```

## Write results

```{r}
st_write(str_clc_case, "./output/str_clc_landuse.gpkg", delete_layer = TRUE)
st_write(str_clc_urban_mask, "./output/str_clc_urbanmask.gpkg", delete_layer = TRUE)

st_write(dor_clc_case, "./output/dor_clc_landuse.gpkg", delete_layer = TRUE)
st_write(dor_clc_urban_mask, "./output/dor_clc_urbanmask.gpkg", delete_layer = TRUE)

st_write(liv_clc_case, "./output/liv_clc_landuse.gpkg", delete_layer = TRUE)
st_write(liv_clc_urban_mask, "./output/liv_clc_urbanmask.gpkg", delete_layer = TRUE)
```
