---
title: "Prepare GHS-Pop"
output: html_document
---

## Load packages

```{r, echo=FALSE}
library(sf)
library(tidyverse)
library(terra)
```

## Set root directory

Set path to your wd in masterfile or directly here. 

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(my_wd))
```

## Load Data

* Case study region
* GHS-POP: 
   + Download Data (choose tiles for case study regions): https://ghsl.jrc.ec.europa.eu/download.php?ds=pop
   + Epoch: 2020
   + Resolution: 100m
   + CRS: Mollweide

```{r}
str_cs <- st_read("./output/str_iso30_commune_dissolved.gpkg")
dor_cs <- st_read("./output/dor_iso30_VWG_dissolved.gpkg")
liv_cs <- st_read("./output/liv_iso30_LSOA_dissolved.gpkg")

r1 <- rast("./input/GHSL/GHS_POP_E2020_GLOBE_R2023A_54009_100_V1_0_R3_C18/GHS_POP_E2020_GLOBE_R2023A_54009_100_V1_0_R3_C18.tif")
r2 <- rast("./input/GHSL/GHS_POP_E2020_GLOBE_R2023A_54009_100_V1_0_R3_C19/GHS_POP_E2020_GLOBE_R2023A_54009_100_V1_0_R3_C19.tif")
r3 <- rast("./input/GHSL/GHS_POP_E2020_GLOBE_R2023A_54009_100_V1_0_R4_C18/GHS_POP_E2020_GLOBE_R2023A_54009_100_V1_0_R4_C18.tif")
r4 <- rast("./input/GHSL/GHS_POP_E2020_GLOBE_R2023A_54009_100_V1_0_R4_C19/GHS_POP_E2020_GLOBE_R2023A_54009_100_V1_0_R4_C19.tif")
```

## Aggregate into one raster and save for cci calculations

```{r}
r_all <- mosaic(r1, r2, r3, r4, fun = "mean")

# Write GeoTIFF, NA values will be written as NoData
writeRaster(r_all, "./output/GHS_POP_E2020_europe.tif", overwrite=TRUE)
```

## Calculate average of neighbouring cells

```{r}
# Define a 3x3 moving window
w <- matrix(1, nrow = 3, ncol = 3)

# Replace 0 values with NA so they are ignored
r_na <- classify(r_all, rbind(c(0, 0, NA)))

# Apply focal mean including the center cell
r_all_mean <- focal(r_na, w = w, fun = mean, na.rm = TRUE)
```

## Derive vectordata with population density and mean population density 

## Crop geotiff to case study regions
```{r}
vectorise_ghspop <- function(region_cs, r_all, r_all_mean) {
  
  # Reproject vector to match raster CRS
  region_cs <- st_transform(region_cs, crs(r_all))
  
  # Crop raster to vector extent and mask - for all
  r_all_crop <- crop(r_all, vect(region_cs))
  r_all_masked <- mask(r_all_crop, vect(region_cs))
  
  # Crop raster to vector extent and mask - for mean values
  r_mean_crop <- crop(r_all_mean, vect(region_cs))
  r_mean_masked <- mask(r_mean_crop, vect(region_cs))
  
  # Vectorize Raster
  
  ## Convert raster to polygons and convert to sf - for all
  vectorized_all <- as.polygons(r_all_masked, aggregate = FALSE) %>%
    st_as_sf(vectorized_all)
  
  ## Convert raster to polygons and convert to sf - for mean values
  vectorized_mean <- as.polygons(r_mean_masked, aggregate = FALSE) %>%
    st_as_sf(vectorized_mean)
  
  # Prepare final dataset
  region_ghs_pop_all <- vectorized_all %>%
    transmute(ghs_pop = GHS_POP_E2020_GLOBE_R2023A_54009_100_V1_0_R3_C18)
  
  region_ghs_pop_all_mean <- vectorized_mean %>%
    transmute(ghs_pop_mean = focal_mean) %>%
    st_drop_geometry()
  
  region_ghs_fin <- region_ghs_pop_all %>%
    bind_cols(region_ghs_pop_all_mean) %>%
    filter(ghs_pop != 0 & ghs_pop_mean != 0)
  
  region_ghs_fin
  
}
```

```{r}
str_ghs_fin <- vectorise_ghspop(str_cs, r_all, r_all_mean)
dor_ghs_fin <- vectorise_ghspop(dor_cs, r_all, r_all_mean)
liv_ghs_fin <- vectorise_ghspop(liv_cs, r_all, r_all_mean)
```

## Write results

```{r}
st_write(str_ghs_fin, "./output/str_ghs_pop.gpkg", delete_layer = TRUE)
st_write(dor_ghs_fin, "./output/dor_ghs_pop.gpkg", delete_layer = TRUE)
st_write(liv_ghs_fin, "./output/liv_ghs_pop.gpkg", delete_layer = TRUE)
```
