---
title: "Prepare data and folder structure for CCI"
output: html_document
---

## Load packages

```{r, echo=FALSE}
library(sf)
library(dplyr)
library(terra)
```

## Set root directory

Set path to your wd in masterfile or directly here. 

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(my_wd))
```

## Create directories needed for processing

Here we create all the directories which we later need in the python processing

```{r}
dir.create("./output/cci")
dir.create("./output/cci/input")
dir.create("./output/cci/output")
dir.create("./output/cci/input/OSM")
dir.create("./output/cci/input/csr")
dir.create("./output/cci/input/csr/buffer")
```

## Load Data

* Case study regions
* GHS-POP prepared raster

```{r}
dor_case <- st_read("./output/dor_iso30_VWG_dissolved.gpkg")
str_case <- st_read("./output/str_iso30_commune_dissolved.gpkg")
liv_case <- st_read("./output/liv_iso30_LSOA_dissolved.gpkg")

r_all <- rast("./output/GHS_POP_E2020_europe.tif")
```

## Prepare population raster  

Exclude 0 values from population raster

```{r}
# Set 0 values to NA
r_all_cci <- r_all
  
r_all_cci[r_all_cci == 0] <- NA

# Set a clean layer name (e.g., "Band1")
names(r_all_cci) <- ""
```

## Create layer with 20.000 m Buffer

This layer is needed in 06c (processing step in Python). It needs to be saved as Shapefile.

```{r}
dor_case <- st_transform(dor_case, crs = 25832)
str_case <- st_transform(str_case, crs = 25832)
liv_case <- st_transform(liv_case, crs = 25832)

dor_buffer <- st_buffer(dor_case, 20000)
str_buffer <- st_buffer(str_case, 20000)
liv_buffer <- st_buffer(liv_case, 20000)
```

## Write results

```{r}
writeRaster(r_all_cci, "./output/cci/input/GHS_POP_E2020_non_0.tif", overwrite=TRUE)

st_write(dor_case, "./output/cci/input/csr/dor.shp", delete_layer = TRUE)
st_write(dor_buffer, "./output/cci/input/csr/buffer/buffer_dor.shp", delete_layer = TRUE)

st_write(str_case, "./output/cci/input/csr/str.shp", delete_layer = TRUE)
st_write(str_buffer, "./output/cci/input/csr/buffer/buffer_str.shp", delete_layer = TRUE)

st_write(liv_case, "./output/cci/input/csr/liv.shp", delete_layer = TRUE)
st_write(liv_buffer, "./output/cci/input/csr/buffer/buffer_liv.shp", delete_layer = TRUE)
```
