---
title: "Join data of all previous steps and derive projects"
output: html_document
---

## Load packages

```{r, echo=FALSE}
library(tidyverse)
library(sf)
```

## Set root directory

Set path to your wd in masterfile or directly here. 

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(my_wd))
```

## Load Data

* Densification dataset: Contains output of building evolution and the resulting densification type; it also includes disappeared buildings, which were there in 2011
* Buildings type: Contains output of building type classification

```{r}
dor_densification <- read_sf("./output/dor_bldg_densetype.gpkg") %>% 
  filter(id_univ != "bldg_182552") # id_bldg = DENW01AL1000CkwP - check in previous files
dor_bldgtype <- read_sf("./output/dor_bldg_2021_03_bldg_type.gpkg") %>% st_drop_geometry()

str_densification <- read_sf("./output/str_bldg_densetype.gpkg") 
str_bldgtype <- read_sf("./output/str_bldg_2022_03_bldg_type.gpkg") %>% st_drop_geometry()

liv_densification <- read_sf("./output/liv_bldg_densetype.gpkg") 
liv_bldgtype <- read_sf("./output/liv_bldg_2021_03_bldg_type.gpkg") %>% st_drop_geometry()
```

# Join data

## Filter disappeared buildings, join building type information and rename variables

```{r}
# Dortmund
# Building type: Results of random forest classification; for Dortmund we only classified buildings which were "res" based on the authorative building use information; some garages and smaller auxiliary buildings were also "res"; they are classified as pred_bldg_type == 99 and should also be added to "non-res"
dor_dat <- dor_densification %>% 
  filter(denstype != "disappeared") %>%  
  transmute(id_bldg = oi_2, denstype) %>% 
  left_join(dor_bldgtype %>% mutate(id_bldg = oi)) %>% 
  transmute(id_bldg,
            bldg_height = Gebaeude_H,
            area_foot, res_non_res, pred_bldg_type, denstype) %>% 
# Assign class 99 to non_res; create sfh_mfh variable and filter variable dens_res_fin  
  mutate(res_non_res = if_else(res_non_res == "non_residential", "non_res", "res"),
         res_non_res = if_else(res_non_res == "res" & pred_bldg_type == 99, "non_res", res_non_res),
         sfh_mfh = if_else(pred_bldg_type == 10, "mfh", "other"),
         sfh_mfh = if_else(pred_bldg_type == 30, "sfh", sfh_mfh),
         sfh_mfh = if_else(is.na(sfh_mfh), "other", sfh_mfh),
         bldg_vol = area_foot*bldg_height) %>% 
  mutate(dens_res_fin = if_else(res_non_res == "res" & denstype %in% c("reconstruction", "infill") , "dens_res", "other"))

# Strasbourg
str_dat <- str_densification %>% 
  filter(denstype != "disappeared") %>%  
  transmute(id_bldg = ID_2, denstype) %>% 
  left_join(str_bldgtype %>% mutate(id_bldg = ID)) %>% 
  transmute(id_bldg,
            bldg_height = Gebaeude_H,
            area_foot, pred_bldg_type, denstype) %>% 
  mutate(res_non_res = if_else(pred_bldg_type %in% c("aux", "non_res"), "non_res", "res"),
         sfh_mfh = if_else(pred_bldg_type == "MFH", "mfh", "other"),
         sfh_mfh = if_else(pred_bldg_type == "SFH", "sfh", sfh_mfh),
         sfh_mfh = if_else(is.na(sfh_mfh), "other", sfh_mfh),
         bldg_vol = area_foot*bldg_height) %>% 
  mutate(dens_res_fin = if_else(res_non_res == "res" & denstype %in% c("reconstruction", "infill") , "dens_res", "other"))

# Liverpool
liv_dat <- liv_densification %>% 
  filter(denstype != "disappeared") %>%  
  transmute(id_bldg = fid_os_2, denstype) %>% 
  left_join(liv_bldgtype %>% mutate(id_bldg = fid_os)) %>% 
  transmute(id_bldg,
            bldg_height = RelH2,
            area_foot, sfh_mfh, denstype) %>% 
  mutate(res_non_res = if_else(is.na(sfh_mfh), "non_res", "res"),
         sfh_mfh = if_else(is.na(sfh_mfh), "other", sfh_mfh),
         bldg_vol = area_foot*bldg_height) %>% 
  mutate(dens_res_fin = if_else(res_non_res == "res" & denstype %in% c("reconstruction", "infill") , "dens_res", "other"))
```

## Function DB Scan clusters for densification projects

```{r}
run_dbscan <- function(region_dat) {
  
  # Filter to densification
  dens <- region_dat %>%
    filter(dens_res_fin == "dens_res")
  
  # Compute Centroids
  centroids <- st_centroid(dens) %>%
    select(id_bldg)
  
  # DBSCAN Clustering
  # Extract coordinates
  coords <- st_coordinates(centroids)
  
  # Run DBSCAN: We set a minimum cluster size of 3 buildings and a cut-off distance of 60 m
  db <- dbscan::dbscan(coords, eps = 60, minPts = 3)
  
  # Add clustering info
  centroids$CLUSTER_ID <- db$cluster
  
  # Derive ID and cluster size
  centroids_id <- centroids %>%
    arrange(CLUSTER_ID) %>%
    mutate(
      cluster_id = if_else(
        CLUSTER_ID == 0,
        paste0("ncl_", row_number()),
        paste0("cl_", CLUSTER_ID)
      ),
      cluster_size = ifelse(CLUSTER_ID == 0, 1, ave(cluster_id, cluster_id, FUN = length))
    ) %>%
    select(-CLUSTER_ID)
  
  ## Join Back to Original Data
  bldg_dens_dbscan <- left_join(region_dat, centroids_id %>% st_drop_geometry(), by = "id_bldg")
  
  # return dataframe
  return(bldg_dens_dbscan)
  
}
```

```{r}
str_dbscan <- run_dbscan(str_dat)
dor_dbscan <- run_dbscan(dor_dat)
liv_dbscan <- run_dbscan(liv_dat)
```

## Function Create Polygons for projects

By buffering the buildings and merging them by cluster_id, we create project polygons. 

```{r}
create_polygons_projects <- function(region_dbscan) {
  
  # fitler to densification
  dat <- region_dbscan  %>%
    filter(dens_res_fin == "dens_res")
  
  # Buffer +20 Meters
  buffered_20 <- st_buffer(dat, dist = 20)
  
  # Dissolve by 'cluster_id'
  dissolved <- buffered_20 %>%
    group_by(cluster_id) %>%
    summarise(geometry = st_union(geom), .groups = "drop")
  
  # Multipart to Singleparts
  # singleparts <- st_cast(dissolved, "POLYGON")
  
  # Delete Holes: Remove holes by keeping only exterior rings
  remove_holes <- function(polygon) {
    if (inherits(polygon, "POLYGON")) {
      st_polygon(list(polygon[[1]]))  # Keep only exterior ring
    } else {
      polygon  # Return as-is for non-polygon geometries
    }
  }
  
  cleaned <- st_geometry(dissolved) %>%
    lapply(remove_holes) %>%
    st_sfc(crs = st_crs(dissolved))
  
  dissolved$geometry <- cleaned
  
  ## Buffer -15 Meters
  projects_dens <- st_buffer(dissolved, dist = -15)
  
  ## Add Area Field
  projects_dens$proj_area <- as.numeric(st_area(projects_dens))
  
  # return dataframe
  return(projects_dens)
  
}
```

```{r}
str_poly_projects <- create_polygons_projects(str_dbscan)
dor_poly_projects <- create_polygons_projects(dor_dbscan)
liv_poly_projects <- create_polygons_projects(liv_dbscan)
```


# Function to join corine land use information, ghs pop and cci to data

```{r}
join_clc_cci_ghs <- function(region, region_dbscan) {
  
  clc <- st_read(paste0("./output/", region, "_clc_landuse.gpkg"))
  ghs <- st_read(paste0("./output/", region, "_ghs_pop.gpkg"))
  cci <- st_read(paste0("./output/", region, "_cci_fin.gpkg"))
  
  # Create centroids of buildings for spatial join
  centroids <- st_centroid(region_dbscan) %>%
    select("id_bldg")
  
  # Make sure that crs is correct and reduce var
  clc <- st_transform(clc, st_crs(centroids)) %>%
    transmute(
      clc_code_simple = code_simple,
      clc_code_nam = code_nam,
      clc_inside = "inside"
    )
  
  ghs <- st_transform(ghs, st_crs(centroids))
  
  cci <- st_transform(cci, st_crs(centroids)) %>%
    transmute(cci = new_m_w_n_idw)
  
  # join attributes by location
  joined <- st_join(centroids, clc, .predicate = st_intersects) %>%
    st_join(ghs, .predicate = st_intersects) %>%
    st_join(cci, .predicate = st_intersects)
  
  # join back to overall data
  bldg_dat_fin <- region_dbscan %>%
    left_join(joined %>% st_drop_geometry())
  
  # return dataframe
  return(bldg_dat_fin)
  
}
```

```{r}
str_bldg_dat_fin <- join_clc_cci_ghs("str", str_dbscan)
dor_bldg_dat_fin <- join_clc_cci_ghs("dor", dor_dbscan)
liv_bldg_dat_fin <- join_clc_cci_ghs("liv", liv_dbscan)
```

## Write results

```{r}
st_write(str_bldg_dat_fin, "./output/str_bldg_2022_04_all_dens.gpkg", delete_layer = TRUE)
st_write(str_poly_projects, "./output/str_projects_01_dens.gpkg", delete_layer = TRUE)

st_write(dor_bldg_dat_fin, "./output/dor_bldg_2021_04_all_dens.gpkg", delete_layer = TRUE)
st_write(dor_poly_projects, "./output/dor_projects_01_dens.gpkg", delete_layer = TRUE)

st_write(liv_bldg_dat_fin, "./output/liv_bldg_2021_04_all_dens.gpkg", delete_layer = TRUE)
st_write(liv_poly_projects, "./output/liv_projects_01_dens.gpkg", delete_layer = TRUE)
```
