---
title: "Derive project type"
output: html_document
---

## Load packages

```{r, echo=FALSE}
library(tidyverse)
library(sf)
```

## Set root directory

Set path to your wd in masterfile or directly here. 

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = normalizePath(my_wd))
```

## Load Data

```{r}
dor_dat <- read_sf("./output/dor_bldg_2021_04_all_dens.gpkg") 
dor_proj <- read_sf("./output/dor_projects_01_dens.gpkg") 

str_dat <- read_sf("./output/str_bldg_2022_04_all_dens.gpkg") 
str_proj <- read_sf("./output/str_projects_01_dens.gpkg") 

liv_dat <- read_sf("./output/liv_bldg_2021_04_all_dens.gpkg") 
liv_proj <- read_sf("./output/liv_projects_01_dens.gpkg") 
```

## For each project (each cluster_id) we derive the following variabels

* Share of building area_foot inside/outside clc
* Predominant building type
* New building volume
* Mean population density

```{r}
derive_projecttype <- function(df_reg, df_prj) {
  
  bldg_type <- df_reg %>%
    st_drop_geometry() %>%
    group_by(cluster_id, sfh_mfh) %>%
    summarise(n_bldg = n(),
              area_foot_sum = sum(area_foot, na.rm = TRUE)) %>%
    pivot_wider(names_from = sfh_mfh,
                values_from = n_bldg:area_foot_sum) %>%
    transmute(cluster_id,
              area_foot_sum_sfh = if_else(is.na(area_foot_sum_sfh), 0, area_foot_sum_sfh),
              area_foot_sum_mfh = if_else(is.na(area_foot_sum_mfh), 0, area_foot_sum_mfh),
              area_foot_all_bldg = area_foot_sum_sfh + area_foot_sum_mfh,
              share_area_foot_sfh = area_foot_sum_sfh / area_foot_all_bldg,
              bldg_type_predominant = if_else(share_area_foot_sfh > 0.5, "sfh", "mfh"))
  
  inside_clc <- df_reg %>%
    st_drop_geometry() %>%
    mutate(clc_inside = if_else(is.na(clc_inside), "outside", clc_inside)) %>%
    group_by(cluster_id, clc_inside) %>%
    summarise(n_bldg = n(),
              area_foot_sum = sum(area_foot, na.rm = TRUE)) %>%
    pivot_wider(names_from = clc_inside,
                values_from = n_bldg:area_foot_sum) %>%
    transmute(cluster_id,
              area_foot_sum_inside = if_else(is.na(area_foot_sum_inside), 0, area_foot_sum_inside),
              area_foot_sum_outside = if_else(is.na(area_foot_sum_outside), 0, area_foot_sum_outside),
              area_foot_all_bldg = area_foot_sum_inside + area_foot_sum_outside,
              share_area_foot_TRUE = area_foot_sum_inside / area_foot_all_bldg,
              proj_clc_inside = if_else(share_area_foot_TRUE > 0.5, "inside", "outside"))
  
  ghs_pop_total <- df_reg %>%
    st_drop_geometry() %>%
    group_by() %>%
    summarise(
      ghs_pop_proj = mean(ghs_pop, na.rm = TRUE),
      ghs_pop_mean_proj = mean(ghs_pop_mean, na.rm = TRUE))
  
  ghs_pop <- df_reg %>%
    st_drop_geometry() %>%
    group_by(cluster_id) %>%
    summarise(
      ghs_pop_proj = mean(ghs_pop, na.rm = TRUE),
      ghs_pop_mean_proj = mean(ghs_pop_mean, na.rm = TRUE)) %>%
    # some projects do have na values, as gridcell in ghs_pop is missing; we substitute them with the mean of all projects
    mutate(ghs_pop_proj = if_else(is.nan(ghs_pop_proj), 43, ghs_pop_proj),
      ghs_pop_mean_proj = if_else(is.nan(ghs_pop_mean_proj), 37, ghs_pop_mean_proj))
  
  new_bldg_vol <- df_reg %>%
    st_drop_geometry() %>%
    group_by(cluster_id) %>%
    summarise(n_bldg = n(),
              bldg_vol_sum = sum(bldg_vol, na.rm = TRUE))
  
  ## Join data and derive project types
  project_types <- df_prj %>%
    left_join(new_bldg_vol) %>%
    left_join(inside_clc %>% select(cluster_id, proj_clc_inside)) %>%
    left_join(ghs_pop) %>%
    left_join(bldg_type %>% select(cluster_id, share_area_foot_sfh, bldg_type_predominant)) %>%
    mutate(proj_type = if_else(n_bldg <= 3 & bldg_type_predominant == "sfh", "S_sfh", "M"),
      proj_type = if_else(n_bldg <= 3 & bldg_vol_sum <= 10000 & bldg_type_predominant == "mfh", "S_mfh", proj_type),
      proj_type = if_else(n_bldg > 30 & bldg_type_predominant == "sfh", "L_sfh", proj_type),
      proj_type = if_else((n_bldg > 10 | bldg_vol_sum > 50000) & bldg_type_predominant == "mfh", "L_mfh", proj_type),
      proj_type = if_else(proj_type == "M" & bldg_type_predominant == "sfh", "M_sfh", proj_type),
      proj_type = if_else(proj_type == "M" & bldg_type_predominant == "mfh", "M_mfh", proj_type))
  
  # return dataframe
  return(project_types)
  
}
```

```{r}
dor_project_types <- derive_projecttype(dor_dat, dor_proj)
str_project_types <- derive_projecttype(str_dat, str_proj)
liv_project_types <- derive_projecttype(liv_dat, liv_proj)
```

## Join back to building data

```{r}
dor_bldg_proj_type <- dor_dat %>% 
  left_join(dor_project_types %>% st_drop_geometry() %>% 
              select(cluster_id, proj_area, proj_clc_inside, proj_type))

str_bldg_proj_type <- str_dat %>% 
  left_join(str_project_types %>% st_drop_geometry() %>% 
              select(cluster_id, proj_area, proj_clc_inside, proj_type))

liv_bldg_proj_type <- liv_dat %>% 
  left_join(liv_project_types %>% st_drop_geometry() %>% 
              select(cluster_id, proj_area, proj_clc_inside, proj_type))
```

## Write results

```{r}
st_write(dor_project_types, "./output/dor_projects_02_projecttypes.gpkg", delete_layer = TRUE)
st_write(dor_bldg_proj_type, "./output/dor_bldg_2021_05_all_dens_fin.gpkg", delete_layer = TRUE)

st_write(str_project_types, "./output/str_projects_02_projecttypes.gpkg", delete_layer = TRUE)
st_write(str_bldg_proj_type, "./output/str_bldg_2021_05_all_dens_fin.gpkg", delete_layer = TRUE)

st_write(liv_project_types, "./output/liv_projects_02_projecttypes.gpkg", delete_layer = TRUE)
st_write(liv_bldg_proj_type, "./output/liv_bldg_2021_05_all_dens_fin.gpkg", delete_layer = TRUE)
```

## Write centroids for QGIS

Optional step for graphic representation in QGIS

```{r}
# region <- c("dor", "str", "liv")
# type <- c("L_mfh", "L_sfh", "M_mfh", "M_sfh", "S_mfh", "S_sfh")
# 
# dir.create("./output/centroids")
# 
# run_write_centroids <- function(region, type) {
#   
#   # Get the appropriate data
#   input_data <- get(paste0(region, "_project_types"))
#   
#   # Calculate centroids and filter
#   centroids <- st_centroid(input_data) %>% 
#     filter(proj_clc_inside == "inside")
#   
#   # Further filter by project type
#   output_data <- centroids %>% filter(proj_type == type)
#   
#   # Write to file
#   st_write(output_data, paste0("./output/centroids/", region, "_projects_centroids_", type, ".gpkg"))
# }
# 
# library(purrr)
# 
# combinations <- expand.grid(region = region, type = type, stringsAsFactors = FALSE)
# pmap(combinations, run_write_centroids)
```






